---
title: "Redes Complexas: Análises Básicas"
author: "Carine Emer"
date: "11/10/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
#######################################################################

Curso de Introdução à Teoria de Redes Ecológicas

Todas as espécies interagem com alguma outra espécie na natureza. Nessa prática os alunos coletarão dados sobre interações mutualísticas na natureza (ex. formiga-planta, dispersor de semente-planta, polinizador-planta) e/ou utilizarão dados próprios ou da literatura para analisar métricas básicas de redes complexas.

No final do curso o aluno será capaz de: 
1) Visualizar uma rede de interações 
2) Testar se rede é a conectada, modular ou aninhada com modelos nulos
3) Identificar a centralidade dos nodos
3) Explorar a estrutura da rede e seu significado biológico

##########################################################################
__________________________________________________________________________
Warning 1: this script works for both binary and weighted networks.

Warning 2: there is no single magic way to draw all kinds of network.
There are several network drawing methods implemented in R packages
and stand-alone software. Study their logic and algorithms,
see some papers in which they were used, think it through,
and only then decide which drawing method to use in your study.(M. Mello)

Warning 3: Null models are an endless discussion in Ecology. The choice of null models depends on your question, but is essential to test the significance of the data. Chose carefully, based on the biological meanings of the randomization process.

Warning 4: always check for updates of packages and R versions, specially if you start to have weird errors.

#Have fun!

#########################################################################


```{r setup, echo=T, include = F}
# Set the working directory
#setwd("paste your working directory path here")
setwd("~/Documents/Documents - Girassol/Network course/Curso Redes UESC 2018/práticas R/data")
knitr::opts_knit$set(root.dir = "~/Documents/Documents - Girassol/Network course/Curso Redes UESC 2018/práticas R/data")
```


```{r load packages, include = F, echo = T}
# load packages
library(bipartite)
library(vegan)
library(reshape2)
library(igraph)
library(networkD3)
library(reshape2)
```

```{r input data, echo=T, include = F}
### input data

### you can use the dataset available or your own data. To do so, just replace "the name within brackets" below for your won.

### Binary networks - qualitative data
poll_b<-read.csv("pollination_NewZealand_bin.csv", head=T,row.names=1)
ants_b<-read.csv("ants_Amazon_bin.csv", head=T,row.names=1)
seed_b<-read.csv("seed_dispersal_AtlanticForest_bin.csv", head=T,row.names=1)


###### Weighted networks - quantitative data
poll<-read.csv("pollination_NewZealand.csv", head=T,row.names=1)
ants<-read.csv("ants_Amazon.csv", head=T,row.names=1)
seed<-read.csv("seed_dispersal_AtlanticForest.csv", head=T,row.names=1)

##### check data
poll
dim(poll)
str(poll)

```

#################################
Drawing networks
#################################

```{r plot networks bipartite}

#### Observe topological differences between binary and weighted networks for the different systems

#### Invaded pollination network from a sub-alpine vegetation, New Zealand (Emer et al. 2016. DOI: 10.1111/ddi.12458)
poll_network<-plotweb(poll,
              ## set overall parameters 
              text.rot=c(90), labsize=1,arrow="down.center",
              y.width.low=0.05,y.width.high=0.05, ybig=1.8, low.y=1.2, high.y=2,
              high.spacing=.01, low.spacing=0.01,
              # shorter labels
              #high.lablength=3, low.lablength=0, 
              ### check the method if you prefer
                   method="cca", 
              ### set colors
                  col.low="green", col.high="yellow", col.interaction = "grey80")

#### Network of ant-myrmecophyte interactions from a dam-fragmented island in Balbina, Central Amazon (Emer et al. 2013. DOI: 10.1111/cobi.12045).

# vectors for all colours can be given, to mark certain species or 
# interactions. Colour vectors are recycled if not of appropriate length - BE CAREFULL
# colors are given according to the number of the cells in the matrix/dataframe (rows x cols). It is counted from right-to-left,up-to-down
# for example, lets color the interaction between Cordia nodosa and Azteca sp5 as blue, creating a vector for that. 

# check size of the network by multiplying the n of rows and columns
dim(ants) # 11*15 = 165
#check the position of the interaction of interest, create the vector
my.col<-c(rep("white",33),rep("blue",1),rep("white",131))                          
myrm_network<-plotweb(ants,text.rot=c(90),            labsize=1,y.width.low=0.05,y.width.high=0.05,arrow="down.center",ybig=1.8, low.y=1.2, high.y=2, high.spacing=.01, low.spacing=0.01,method="cca",
                      col.interaction = my.col)


#### Avian seed-dispersal network from Ilha Anchieta, Atlantic Forest (Bello et al. 2017. DOI doi/10.1002/ecy.1818/).
seed_network<-plotweb(seed,text.rot=c(90),labsize=1,y.width.low=0.05,y.width.high=0.05,
        arrow="down.center",ybig=1.8, low.y=1.2, high.y=2, high.spacing=.01, low.spacing=0.01,
        method="cca", ### check the method if you prefer
        col.low="green", col.high="yellow", col.interaction = "grey80")

#export the image
#png('seed_network.png')
       
#dev.off()
# Tip: There are many other parameters that can be set to customize the drawing.
# Explore them!

### for examples of tri-trophic interactions and how to add abundances in the axis see "script_visual_Gruber_Safariland.R" (https://rdrr.io/cran/bipartite/man/plotweb.html)
```

```{r plot networks igraph}
### some more complex but flexible ways of drawing networks
# ??str.igraph
### code needs update to differentiate plants and animals
m<- read.csv("ants_Amazon.csv", head=T)
clos<-read.csv("closeness_ants.csv", head=T)

# transforms the column to a list of character strings
  clos$species <- as.character(clos$species) 
# transforms the column to a numeric vector
  clos$closeness <- as.character(clos$closeness) 
  clos$guild<-as.character(clos$guild) 

## transform to edgelist
m1<-melt(m) 
## select only interactions =1
m2 <- subset(m1, value>0) 

# sets the graph framework
  g=graph.data.frame(m2,directed=FALSE) 
#Check the main properties of the igraph object
V(g)
E(g)
# ordering clos so that the order of species list matches the order of species in graph g
  clos <- clos[order(match(clos$species, V(g)$name)),] 
  clos$closeness <- as.numeric(clos$closeness)

# to a given value of closeness you associate a colour
  #color <- colorRampPalette(c('white','blue')) # colour palette
# the colour vector for each species
  closeness.col <- heat.colors(25)[as.numeric(cut(-clos$closeness, breaks = 25))] 


## to create a vector for guild
V(g)$guild=as.character(clos$guild)
V(g)$shape=V(g)$guild
V(g)$shape[V(g)$guild=="ant"]="circle"
V(g)$shape[V(g)$guild=="plant"]="square"

plot(g,
      # Set the drawing mode
      # This package contains several drawing methods; try them!
     layout=layout_nicely,
       # Set title
     main='Ant-myrmecophyte network',
       # Set node sizes
     vertex.size=6,
     # Set node attributes
     vertex.shape=V(g)$shape,
     vertex.size = 12,
     vertex.color = closeness.col,
     # Set link colors
           edge.color = "lightblue",
      # Set link curvature from 0 to 1
           edge.curved=0.3,
      # Set nodes labels
     vertex.label.dist=200,
     vertex.label.color='black',
     vertex.label.font=1,
     vertex.label=V(g)$name,
     vertex.label.cex=0.5
           )
#export the image
#png('antplant_network.png')

############################## BIPARTITE NETWORKS W/ IGRAPH #########
V(g)$type <- V(g)$name %in% m2[,1]
bipartite.projection(g)
V(g)$x <- V(g)
V(g)$y <- E(g)
col <- c("steelblue", "orange")
shape <- c("circle", "square")

V(g)$shape <- shape[as.numeric(V(g)$type) + 1]
V(g)$color <- c('red', 'blue', 'green', 'steelblue', 'steelblue', 'steelblue', 'steelblue')
E(g)$color <- 'gray'
plot(g)

  
```

```{r plot networks network3D}
n3D<-simpleNetwork(m2)
n3D
#saveNetwork(n3D, file = "ants_network3D.html") ### save network as html
```

#################################
Network Structure and null models 
#################################

```{r metrics network structure}

### Observe the differences in some metrics for the analyses of netwoks. Run the metrics for other systems, compare the topology among them. What does it mean?

#### Explore possibilities of different null models in the different packages, for example:
??oecosimu
??vegan::commsim
??nullmodel

############## Degree distribution##########################
#??degreedistr ## check the help for some limitations of this method

poll.dd <- degreedistr(poll)
poll.dd
#Dark grey: exp
#Medium grey: power law
#Light grey: truncated power law

############################## CONNECTANCE #####################
obs_c<-networklevel(ants, index="connectance")
# Set Null Model
null_r2d<-oecosimu(ants,networklevel, "r2d", index="connectance",nsim=100)
null_r2d
densityplot(permustats(null_r2d))


############################ NESTEDNESS ########################## 
nested_ants<-networklevel(ants, index = "weighted NODF")
nested_ants

# Set Null Model
  obs_nested<-nested_ants
  null1<-oecosimu(ants,nestednodf,"quasiswap_count",nsimul = 100, order=FALSE)
  null1
  densityplot(permustats(null1))

#### Using the classical nullmodels for bipartite
## method 3 = vaznull, explore the other possibilities and discuss when to chose one or another
  null <-nullmodel(ants, N=100, method=3) 
  null1 <-sapply (X=null, FUN=nested, method="NODF2") 
  means_null1 <- apply (X=cbind(null1),MARGIN=2, FUN=mean, na.rm=T)  
  sd.means_null1 <- apply(X=cbind(null1), MARGIN=2, FUN=sd, na.rm=T)
  z_score<-(obs_nested-means_null1)/sd.means_null1
  z_score
  p_value<-sum(null1>= obs_nested)/1000 # valor de p
  p_value

plot(density(null1), lwd=2, xlim=c(0, 40))
abline(v=obs_nested, col="red", lwd=1)

############################ MODULARITY ########################## 
m_ants<-computeModules(ants)
m_ants
# Check the components of each module
  printoutModuleInformation(m_ants)
  plotModuleWeb(m_ants)

# Set Null Model
  nulls <- nullmodel(ants, N=100, method=1) 
  modules.nulls <- sapply(nulls, computeModules)
  like.nulls <- sapply(modules.nulls, function(x) x@likelihood)
  z <- (m_ants@likelihood - mean(like.nulls))/sd(like.nulls)
  p <- 2*pnorm(-abs(z))
  plot(density(like.nulls), lwd=2, xlim=c(0, .8))
  abline(v=m_ants@likelihood, col="red", lwd=1)
 
  

#save.image("pratica_redes.RData")
```

#################################
Network centrality
#################################
```{r centrality}
#### identify the degree, closess and betweenness of your species

## To test different indexes, you just have to change the index parameter. Level defines whether the index is calculated for the rows (lower) or columns (higher).

clos_ants<-specieslevel(ants,index = "closeness", level= "higher")

### Explore other indexes by yourself

########## Estimating hubs and connectors according to Olensen et al. 2007, PNAS.
cz <- czvalues(m_ants)
plot(cz[[1]], cz[[2]], pch=16, xlab="c, participation coefficient P", ylab="z, within-module degree", cex=0.8, xlim=c(-1,2.5), las=1, ylim= c(-1.5,4))
abline(v=0.62) # threshold of Olesen et al. 2007
abline(h=2.5)   # dito
text(cz[[1]], cz[[2]], names(cz[[1]]), pos=4, cex=0.7)




#### save the data
#write.csv(clos_ants, "closeness_ants_higher.csv")



### Save your results, so you dont have to run everything again once you want to update your code.

save.image("pratica_redes.RData")
#dev.off()
```


```{r robustness}
######################## NETWORK ROBUSTNESS

#This index is based on:
#Burgos, E., H. Ceva, R.P.J. Perazzo, M. Devoto, D. Medan, M. Zimmermann, and A. Maria Delbue (2007) Why nestedness in mutualistic networks? Journal of Theoretical Biology 249, 307ñ313
#Slope.bipartite È baseado em 
#Memmott, J., Waser, N. M. and Price, M. V. (2004) Tolerance of pollination networks to species extinctions. Proceedings of the Royal Society B 271, 2605ñ2611
## However, there are many implications on using it, mainly because it does consider rewiring. Check the literature for updates.


#Removing species randomly
par(mfrow=c(1,2))
#retirada aleatória
ants.extlower.r <- second.extinct(ants, participant="lower", 
                                        method="random", nrep=100, details=FALSE)
robustness(ants.extlower.r)
#png(filename="dispersao1extlower_random.png", width = 500, height = 500)
slope.bipartite(ants.extlower.r, pch=19, cex=.5)
#dev.off()

??str.igraph

#Removing species according to their degree
ants.extlower.d <- second.extinct(ants, participant="lower", method="degree", nrep=100, details=FALSE)
robustness(ants.extlower.d)
#png(filename="dispersao1extlower_degree.png", width = 500, height = 500)
slope.bipartite(ants.extlower.d, pch=19, cex=0.5)
#dev.off()



```


#########################################################################
Now you are ready to explore other parameters of the networks and play with your drawings. Networks are a world of possibilities. Be creative and you will succeed!! 
#########################################################################
